\sout{Rare-event sampling applied to the simulation of extreme mechanical efforts exerted by a turbulent flow on a bluff body} \ZZ{Numerical study of extreme mechanical efforts exerted by a turbulent flow on a bluff body by direct and rare-event sampling techniques}

### 2020-02-06 jfm/draft_p2_jfm.tex
Implement Emmanuel's modifications for section 2 and 3. Move all 3 sections into separate .tex files

\section{Description of the numerical case study}
\label{sec:test_flow}

\begin{figure}
	\centering
	\includegraphics[width=\linewidth]{illustr_ecoulement/illustr_ecoulement}
	\caption{Our case study is a grid-generated turbulent flow impinging onto a fixed squared obstacle (of size $R$) located at the centre of a channel in two dimensions. The flow is artificially damped near the end of the channel. In the developed flow, turbulent eddies have typically the size of the square, which results in strong fluctuations of mechanical efforts acting on the square. The vorticity is displayed with an arbitrary colour map from blue (negative values) to red (positive values).}
	\label{fig:illustr_ecoulement}
\end{figure}

% introduce the flow
%
The drag exerted by a grid-generated turbulent flow onto a fixed squared obstacle is considered as a representative case study (see Fig.~\ref{fig:illustr_ecoulement}).
% why this flow?
%
Although real-world applications would eventually imply three-dimensional dynamics, a simplified two-dimensional setting has been chosen here to reduce the computational cost and allow for a systematic study.
%
We believe that this system embeds the characteristic features that makes \sout{the application of rare-event algorithms} \EL{this study} both relevant and challenging for fluid-structure-interaction problems.
%Namely, spatio-temporal chaos and the emergence of large-amplitude mechanical efforts on the structure.
%
% rapid description of the flow
Turbulent eddies generated in the near-wake of the grid are carried downstream.
They interact with each other and grow in size as expected for two-dimensional turbulent dynamics.
The dimension of the grid is such that the size of the eddies that hit the square is comparable to its size, resulting in strong fluctuations of the drag acting on the square.
%
%Finally, the obstacle does not deform or move either.
%
%Through this simplified setting, our motivation is primarily to evaluate the operability of sampling techniques to capture extreme events with a significant run-time savings.
%

% lattice Boltzmann method
%
The flow dynamics is integrated by the \ac{lbm} in our numerical simulations.
While traditional methods in computational fluid dynamics rely on a discretization of the Navier-Stokes equations, the \ac{lbm} considers the fluid at a mesoscopic level.
Capturing the dynamics of collections of fluid particles moving and colliding on a lattice is here preferred to solving non-linear PDEs.
%This seems crazy, however, most details at the mesoscopic level play actually no role at the macroscopic level. Therefore, the LB algorithm may be viewed as a minimal kinetic scheme compliant to the fluid dynamics at the macroscopic level.
Further details about the \ac{lbm} are given in Appendix \ref{app:lbm} and references therein.
In our context, this numerical method has been chosen principally for its computational manageability and efficiency.

% geometry
%
The simulated flow develops in a long plane channel of dimension $513 \times 129$ mesh points. The square obstacle has size $R=16$ (in mesh unit) and is located at the centre of the channel. The spacing and bar height of the entrance grid are both equal to $R/2$ (see Fig.~\ref{fig:illustr_ecoulement}).
% boundary conditions
%
No-slip boundary conditions are enforced on top and bottom walls of the channel and on the surface of the obstacle by using an halfway bounce-back procedure~\citep{lbm_book}.
%
Upstream of the grid, a constant parabolic velocity profile and a constant mass density (equal to unity) are imposed as an inlet condition.
The centerline velocity is $0.05$ in lattice units, \textit{i.e.} normalised by $\Delta x$ and $\Delta t$ referring to the lattice resolution and the time-step respectively. The initial distributions are imposed at equilibrium (see Appendix).
In the bulk, the viscosity is adjusted so that grid turbulence is generated with Reynolds number $\mathrm{Re_{grid}}=1200$. The reference Mach number is equal to $0.06$ in agreement with the assumption of weak compressibility of the \ac{lbm}.
Near the end of the channel, the flow is progressively damped within a \textit{sponge layer} where the viscosity is artificially enhanced.
Finally, the outlet boundary condition relies on a second-order extrapolation of the velocity and mass density.
The extrapolated distributions are evaluated through a regularization procedure relying on a finite difference estimation of the local stress tensor, as introduced in \cite{latt2008straight}.

\subsection{The drag force}
\label{sec:drag_force}

\begin{figure}
	\centering
	\includegraphics[width=\linewidth]{ecoulement_typique/ecoulement_typique.png}
	\caption{Snapshots of the vorticity related to typical drag fluctuations (within one standard deviation) over a time interval of length \EL{$\tau_c \simeq 4T_0$}; $\tau_c$ will later be identified as the correlation time of the drag signal.
		The vorticity is given in lattice units.}
	\label{fig:typical_vorticity}
\end{figure}

%% LB parameters  %

% drag signal
%
The incoming turbulent flow exerts fluctuating mechanical efforts onto the squared obstacle.
The \textit{drag} is defined as the resulting force in the streamwise $x$-direction. Formally
\begin{equation}
\label{eq:drag_definition}
f_d(t) = \int_{\mathcal{S}} \boldsymbol{\tau}_{x \beta}(\mathbf{x},t) ~ \mathrm{d}{\mathcal{S}}_\beta(\mathbf{x}),
\end{equation}
where $\mathcal{S}$ is the surface of the obstacle and $\boldsymbol{\tau}$ denotes the stress tensor (see Appendix \ref{app:lbm})).
Here, the viscous stress makes a negligible contribution to the drag. The latter therefore results mostly from pressure forces.
% which are closely related to the distribution of velocity gradients in the vicinity of the obstacle (in the nearly-incompressible limit).
%
Since the pressure on the top and bottom sides of the square applies in the normal direction, they do not contribute to the drag.
As a consequence, the drag can eventually be expressed as the difference
\begin{equation}
\label{eq:drag_approx}
f_d(t) = p_{fb}(t) - p_{base}(t)
\end{equation}
between the pressure integrated over the upstream side of the obstacle or \textit{forebody}, $p_{fb}(t)$, and the downstream side or \textit{base} $p_{base}(t)$.
Pressure fluctuations are related to the dynamics of the vorticity \EL{field}.
Regions of strong vorticity correspond to strong local pressure gradients, \emph{e.g.} as demonstrated analytically with a Rankine vortex.
%
% Considering a disc of constant vorticity $\omega$ with radius $a$ (Rankine vortex model), the local pressure gradient between the vortex core and a point infinitely far grows quadratically with the vorticity.
%

% Definition of turnover time
%
The typical timescale (turnover time) of drag fluctuations can be estimated from dimensional analysis as
\begin{equation}
\label{eq:turnover_time}
\EL{T_0 = \frac{R}{U}},
\end{equation}
where $R$ is the size of the square and $U$ is the averaged velocity in the channel.
%The viscosity is here discarded since viscous stress is negligible as compared to pressure forces.
%
% typical fluctuation of the drag
%
Fig.~\ref{fig:typical_vorticity} displays the typical evolution of the vorticity field around the obstacle over a few turnover times.
Because the vorticity generated along the forebody is swept away by the mean flow, the pressure field in the vicinity of the base is only slightly perturbed.

\subsection{The drag as a random process
	%: Probability Density Function and correlation in time
}
\label{sec:pdfs}

\begin{figure}
	\centering
	\includegraphics[width=0.7\linewidth]{typical_drag_signal/typical_drag_signal.png}
	\caption{Temporal evolution of the drag (in lattice units) acting on the square under the action of the impinging turbulent flow. The time is normalised by the turnover time related the mean-flow velocity and the size of the obstacle, \emph{i.e.} \EL{$T_0=R/U$}.}
	\label{fig:typical_drag_signal}
\end{figure}

% drag signal
%
% Description du signal de trainee typique
Figure~\ref{fig:typical_drag_signal} shows the time signal of the drag acting on the square, $f_d(t)$, over five hundred turnover times.
The signal appears unpredictable in details and exhibits repeated bursts of high amplitude that deviate significantly from the averaged value.
Therefore, it is natural to model the drag as a (scalar) random process.

% drag statistics
% pdf
Drag fluctuations have been sampled along a simulation of duration \EL{$T_{tot} = 4\times 10^6~T_0$}.
This long simulation will be referred to as the \textit{control run} in the following.
It has been made possible by the relative simplicity of the investigated flow and the computational efficiency of the lattice Boltzmann method.
The \ac{pdf} of drag fluctuations is shown in Fig.~\ref{fig:pdf_drag_a}.
It deviates from a normal law and shows an exponential tail for large positive fluctuations, \textit{i.e.}  \EL{${\mathbb{P}}(f_d) \propto e^{-\ell f_d}$}.
%
Fig.~\ref{fig:pdf_drag_a} also displays the \ac{pdf} of drag fluctuations acting on a control surface corresponding to the periphery of the obstacle but in the absence of the obstacle.
%
In that case, the \ac{pdf} is quasi-symmetric and does not display exponential tails. This shows that the asymmetry of the \ac{pdf} and the development of a positive exponential tail are closely related to the no-slip condition on the obstacle boundary.
\begin{figure}
	\centering
	\subfloat[\ac{pdf} of (zero-mean) drag fluctuations]
	{\label{fig:pdf_drag_a}
		\includegraphics[width=.45\linewidth]{./PDF_drag/PDF_drag.png}}
	\subfloat[Autocorrelation of drag fluctuations]
	{\label{fig:pdf_drag_b}
		\includegraphics[width=.45\linewidth]{./autocorr_drag/autocorr_drag.png}}
	\caption{\textbf{(a)} \ac{pdf} of (zero-mean) drag fluctuations $\tilde f_d \equiv f_d - \bar{f_d}$ where $\bar{f_d}$ denotes the time-averaged value. The drag is evaluated both in the presence (red) and in the absence (blue) of the obstacle. %Note that the amplitudes have not been normalized.
		\textbf{(b)} Autocorrelation function of the drag defined as $C(\tau) = \overline{ \tilde f_d(t+\tau)\tilde f_d(t)} ~/~ \overline{{\tilde f_d}^2}$. The correlation time \EL{$\tau_c\simeq 4 T_0$} is defined by $C(\tau_c)=0$.
		%$\tau_c$ may therefore be considered as the correlation time of the drag signal.
	}
	\label{fig:pdf_drag}
\end{figure}

% drag statistics
% correlation time
Lastly, the autocorrelation function of the drag $C(\tau)$ is shown in Fig.~\ref{fig:pdf_drag_b}. It is found that drag fluctuations are correlated over a time interval \EL{$\tau_c \simeq 4T_0$}. One can then argue that \EL{the drag loses its memory}  over a time scale corresponding to the sweeping of a few eddies past the obstacle.
%
This observation is important for the application of rare-event algorithms as it will be discussed in section~\ref{sec:rare_events_algorithms}.
%
In the following, $\tau_c$ will be referred to as the \textit{correlation time} of the drag process.
The ratio \EL{$T_0 / \tau_c$} may be viewed as a {Strouhal number}. The value $St=0.25$ is consistent with common observations for flows past blunt structures at comparable Reynolds numbers \citep{rodi1998}.

\section{Extreme fluctuations of the drag by means of direct sampling}
\label{sec:direct_sampling}

\begin{figure}
	\centering
	\includegraphics[width=\linewidth]{illustrate_return_time/illustrate_return_time}
	\caption{\label{fig:illustrate_return_time} {The \textit{return time} $r(a)$  is the averaged waiting time between the occurrence of peak fluctuations of amplitude larger than $a$.
			One observes that $r(a) >> \tau_c$ (correlation time) if $a$ is sufficiently large. The selected peak fluctuations are therefore well separated.}
	}
\end{figure}

\begin{figure}
	\centering
	\includegraphics[width=.6\linewidth]{return_time/return_time.png}
	\caption{Amplitude of drag fluctuations as a \EL{function of} the corresponding return time. $\tilde{f}_d$ denotes the drag with zero mean, \textit{i.e.} $\tilde{f}_d = f_d - \overline{f_d}$.
	}
	\label{fig:return_time_instant}
\end{figure}

% return time given from direct sampling
%
The phenomenology of extreme fluctuations of the drag is first investigated through brute-force direct sampling applied to the control run.
Direct sampling is here used as opposed to approaches involving rare-events algorithms discussed in section~\ref{sec:rare_events_algorithms}.
It will provide a trustworthy baseline for the validation of rare-events algorithms.

The waiting times $\tau$ are defined as the time between two consecutive occurrences of peak fluctuations with amplitude $f_d \geq a$, as illustrated in Fig.~\ref{fig:illustrate_return_time}.
The mixing time $\tau_m$ is the time needed for the dynamics to \EL{lose the memory} of its initial condition.
As soon as the typical waiting times are much larger than the mixing time $\tau_m$, the occurrences of such events follow a Poisson process and the distribution of the waiting times is exponential, \emph{i.e.} $P(\tau)=\lambda(a)\exp(-\lambda(a)\tau)$ where $r(a)=1/\lambda(a)$ is the averaged waiting time \cite{lestang_computing_2018}; $r(a)$ is called the {\it return time} of the level $a$.
For systems without multi-stability, it is common for the mixing time $\tau_m$ to be of the order of the correlation time $\tau_c$.

How rare is a fluctuation $a$ is quantified by the return time $r(a)$.
We can define extreme drag fluctuations as \textit{rare events} in the sense that the return time is much larger than the correlation time, \emph{i.e.} $r(a) \gg \tau_c$.
%
%
%The shape of the return time $r$ as a function of the return level $a$ is system dependent.
If one assumes that  $r(a) = t(a)~/~\mathbb{P}(f_d\geq a)$   where the time scale $t(a)$ is of order $\tau_c$ and varies much more slowly with $a$ than ${\mathbb{P}(f_d\geq a)}$,
% As the PDF ${\mathbb{P}(f_d\geq a)}$ was shown to be well approximated by an exponential (see figure Fig.~\ref{fig:pdf_drag})
one might expect that
\begin{equation}
\label{eq:return_time}
\EL{r(a) \underset{a\to\infty}{\propto} \exp(la)}
\end{equation}
where $l$ is the rate describing the positive tail of the \ac{pdf} of the drag (shown in Fig.~\ref{fig:pdf_drag}).
Fig.~\ref{fig:return_time_instant} shows the evolution of the return time $r(a)$ with the amplitude of fluctuation $a$, computed from {direct sampling} of the drag signal $f_d(t)$~\citep{lestang_computing_2018}.
Consistently, it is found that the return time $r(a)$ is well approximated by an exponential for large levels $a$. Let us also point out some deviation from the exponential law at the largest levels, which are probably the consequence of under-sampling.

% typically for $a \gtrsim 8 \sigma$ with $\sigma$ being the standard deviation.

\subsection{Extracting extreme drag fluctuations from a very long timeseries}
\label{sec:extreme_extraction}

% intro
%
We have extracted the fluctuations of the drag with a return time $r(a)$ greater than  $10^4\tau_c$ from the control time-series $\{f_d(t)\}_{0 \leq t \leq T_{tot}}$.
This set will be considered as representative of \emph{extreme events} in the upcoming study.
The choice of this particular threshold has been driven by the need to collect enough events with large amplitude and possibly identify generic features.
%
According to Fig.~\ref{fig:return_time_instant}, the related amplitude $a$ is found equal to $7.6~\sigma$ with $\sigma$ being the standard deviation of the drag process.
Precisely, 104 independent fluctuations with $f_d(t) \geq 7.6\sigma$ have been identified. Each fluctuation is characterized by its maximal value, $f_d^{\star}$, and the time, $t^{\star}$, at which this maximum is reached.
%
In the following, the phenomenology of extreme drag fluctuations will be examined on the basis of this set of events.

\subsection{\EL{Extreme fluctuations of the instantaneous drag}}
\label{sec:instantaneous_drag}

\subsubsection{Contribution of forebody and base pressure fluctuations to the overall drag fluctuation}
\label{sec:forebody_and_base_contribution}

\begin{figure}
	\centering
	\includegraphics[width=.8\linewidth]{pressure_ratio/pressure_ratio.png}
	\caption{\label{fig:pressure_ratio} Relative contributions of the forebody and base pressure variations to extreme amplitudes of the drag. An extreme event corresponds to an amplitude $\tilde f^{\star}_d$ and a unique pair  ($\tilde{p}^{\star}_{base}$,~$\tilde{p}^{\star}_{fb}$).
		As the amplitude of the drag increases, the relative variation of the base pressure also increases. \EL{The dotted lines represent a linear least-squares fitting of the relative pressure contributions as a function of the (normalized) drag.}
	}
\end{figure}

% teasing
%
In section~\ref{sec:test_flow}, it was pointed out that typical drag fluctuations originate mostly from the variation of the forebody pressure, \textit{i.e.} from the upstream turbulent flow.
We shall see that the situation is different in the case of {extreme} drag fluctuations.

% relative contribution of forebody and base pressure
%
Let $(t^{\star}, f_d^{\star})$ refer to an extreme-drag event.
The (zero-mean) fluctuation $\tilde{f}_d^{\star} = f_d^{\star} - \overline{f_d}$ can be  decomposed into
\begin{equation}
\tilde{f}_d^{\star} = \Delta p_{fb}^{\star} - \Delta p_{base}^{\star}
\end{equation}
where $\Delta p_{fb}^{\star}$ and $\Delta p_{base}^{\star}$ denote the variations of the forebody and base pressure, respectively.
%
Fig.~\ref{fig:pressure_ratio} displays the relative contributions
$\Delta p_{fb}^{\star}/\tilde{f}_d^{\star}$ and $-\Delta p_{base}^{\star}/\tilde{f}_d^{\star}$ to the overall drag fluctuation $\tilde f_d^{\star}$.
%
It is found that the base pressure variation contributes typically to $80\%$ of the overall drag fluctuation.
Therefore, extreme amplitudes of the drag are dominated by the variation of the pressure in the vicinity of the base of the obstacle, \emph{i.e.} downstream of the obstacle.
Furthermore, Fig.~\ref{fig:pressure_ratio} suggests that the larger the fluctuation, the more important is the relative contribution of the base pressure.
%relatively to the forebody pressure.

\subsubsection{Fluid dynamics related to extreme drag fluctuations}
\label{sec:dynamical_aspects}

\begin{figure}
	\centering
	\includegraphics[width=.7\linewidth]{timeseries_extremes/timeseries_extremes.png}
	\caption{\label{fig:timeseries_extremes} Ensemble average of drag signals centred around extreme fluctuations occurring  at $t=t^{\star}$. The blue line shows the mean profile whereas the shaded area indicates variations (around the mean profile) within one standard deviation. Extreme-drag events exhibit a typical lifetime of one correlation time $\tau_c$. The profile is slightly skewed indicating that the step up is slower than the return to typical values.}
\end{figure}

% mean profile around burst
%
The focus is now on the flow scenarii that yield extreme values of the drag.
Fig.~\ref{fig:timeseries_extremes} displays the mean profile (in time) of the drag signal around extreme events. A peaked profile is observed with a width roughly corresponding to one correlation time $\tau_c$. This shows that the duration of extreme events corresponds typically to the sweeping time of the flow past the obstacle.
%Starting from typical values, extreme drags are typically reached in less than a correlation time.
Interestingly, the profile is also slightly skewed indicating that the step up of the drag is slower than the return to typical values past the peak value.
%This is reminiscent of time-irreversibility in turbulent dynamics.
\EL{This asymmetry (under time reversal) is closely linked to the symmetry breaking in what happens before and after the obstacle.}
%Moreover, such very high drag levels do not persist over time.
To better understand the flow scenarii leading to these events, the vorticity fields around the obstacle are now examined.

\begin{figure}
	\centering
	\includegraphics[width=\linewidth]{illustr_extrms_vorticity/illustr_extrms_vorticity.png}
	\caption{\label{fig:top_4_events_vorticity} Vorticity field (in lattice units) around the obstacle at $t=t^{\star}$ for the highest drag amplitudes recorded in the control run.
	}
\end{figure}

\begin{figure}
	\centering
	\includegraphics[width=.5\linewidth]{illustr_density_streamlines/illustr_density_streamlines.png}
	\caption{\label{fig:density+streamlines} Pressure field (in lattice units) and velocity streamlines at $t=t^{\star}$. In the near wake of the square, a (blocking) vortex blocks an intense vortex against the base of the obstacle.}
\end{figure}

\begin{figure}
	\centering
	\includegraphics[width=.8\linewidth]{dynamics_extremes/dynamics_extremes.png}
	\caption{\label{fig:vorticity_dynamics} Snapshots of the vorticity field (in lattice units)  around $t=t^{\star}$.}
\end{figure}
% flow scenario
%
Fig.~\ref{fig:top_4_events_vorticity} displays the vorticity field (in lattice units) around the obstacle for the highest amplitudes of the drag during the control run.
%
In each case, an intense vortical structure is visible near the base of the obstacle.
The vorticity level of this structure is typically twice the amplitude of typical vorticity fluctuations observed in Fig.~\ref{fig:typical_vorticity}.
%+
The formation of this vortex originates from an intense negative (or positive) vorticity layer at the top (or bottom) boundary of the obstacle.
%
This high vorticity is responsible for a significant pressure drop at the base of the obstacle and therefore a strong drag.
In contrast, nothing special happens near the forebody of the obstacle during extreme-drag events.

The high pressure drop near the base of the obstacle appears to be closely related to the presence of a strong vortex blocked against the base.
As illustrated in Fig.~\ref{fig:density+streamlines}, this blockage is enforced by the presence of opposite vorticity in the near wake, which holds the vortex against the base of the obstacle and prevents it from being swept away for a while.
%
This scenario is better evidenced by Fig.~\ref{fig:vorticity_dynamics}, where the time history of the vorticity field around $t=t^\star$ for the same event is shown.
%
Before the occurrence of the extreme event, positive vorticity originating from the bottom boundary layer develops in the near wake of the square. This positive vorticity  prevents the shedding of negative vorticity and enforces the development of a intense vortex against the base of the square.
As the blocking vortex is in turn advected downstream, the vortex against the base is released.
%
Consistently, one can argue that the typical duration of this scenario is related to the sweeping time of the flow past the obstacle, and is therefore of the order of $\tau_c$.
This is in full agreement with the typical duration obtained from statistical consideration on the mean profile of large-drag fluctuations in Fig.~\ref{fig:timeseries_extremes}.
This scenario is generic and has been observed for most extreme events sampled in the control run.

%We found that $80\%$ of the extreme events sampled from the control timeseries can be related to very similar dynamics.

\begin{figure}
	\centering
	\includegraphics[width=.7\linewidth]{shear_asof_drag/shear_asof_drag}
	\caption{\label{fig:shear_asof_drag} Evolution of the (integrated) shear along the top or bottom sides of the obstacle as a function of the drag for $t^{\star}-2\tau_c \leq t \leq t^{\star}-2\tau_c$. Each trajectory corresponds to a single event. The blue line is the mean path averaged over the set of extreme events sampled in the control run.}
\end{figure}

Since the occurrence of large drag amplitudes arises from the production of vorticity along the top or bottom side of the square, it is proposed to characterize the dynamics of extreme events by their trajectory in the parameter space $(f_d(t), \bar{\gamma}(t))$ where $\bar{\gamma}(t)$ is the averaged shear along the top or bottom boundary of the square:
\begin{equation}
\label{eq:avg_shear_def}
\overline{\gamma} = \frac{1}{R} \int_{\mathcal{S}_\parallel} \frac{\partial u(\mathbf{x})}{\partial y}\mathrm{d}\mathbf{x},
\end{equation}
where $R$ denotes the size of the square, $u$ is the streamwise component of the velocity field and $\mathcal{S}_\parallel$ is the surface of either the top or the bottom boundary.
%
% Figure~\ref{fig:type_1_and_type_2_a} displays the label of the 104 sampled events, sorted according to the amplitude of the corresponding fluctuation.
% It illustrates that the dynamics corresponding to the large majority of events also corresponds to the events with the highest fluctuation amplitude.
%
Fig.~\ref{fig:shear_asof_drag} shows $\overline{\gamma}(t)$ as a function of the instantaneous drag $f_d$(t) for $t^{\star}-2\tau_c \leq t \leq  t^{\star}+2\tau_c$ for the 104 sampled extreme events.
Before and after the extremal fluctuation, \textit{i.e.} for $t^{\star}-2\tau_c \leq t \leq t^{\star}-\tau_c$ and $t^{\star}+\tau_c \leq t \leq t^{\star}+2\tau_c$, paths wander in a region related to typical values of both $\overline{\gamma}$ and $f_d$.
On the contrary, the drag abruptly varies for $t^{\star}-\tau_c \leq t \leq t^{\star}+\tau_c$ near the extremal amplitude.
%Paths in the $(f_d, \bar{\gamma})$ space display excursions to atypical values for both $\bar{\gamma}$ and $f_d$.
These excursions always go clockwise, that is, $\overline{\gamma}$ attains its maximum value before $f_d$ does.
This is consistent with an increase of $\overline{\gamma}$ acting as a precursor for extreme drag fluctuations.
In this representation, we also observe that the path related to the increase of the drag is longer than the path related to the return to typical values.

\subsection{Extreme fluctuations of the time-averaged drag }
\label{sec:time_avg}

\begin{figure}
	\centering
	\includegraphics[width=\linewidth]{timeseries_extrms_AVG/timeseries_extrms_AVG}
	\caption{Instantaneous drag signals $f_d(t)$ corresponding to the highest fluctuations of the averaged  drag $F_T$ with a time window $T = 10 \tau_c$;  $\sigma$ and $\sigma_T$ denote the standard deviations of the instantaneous and averaged drag, respectively.}
	\label{fig:extreme_avg}
\end{figure}

% motivation
%
%In section~\ref{sec:instantaneous_drag}, we discussed the phenomenology of rare events corresponding to extremely high values of the drag acting on the square obstacle mounted in the flow described in section~\ref{sec:test_flow}. In particular, it was pointed out that such extreme drag fluctuations have a lifetime of, roughly, one correlation time $\tau_c$.

%We discussed previously the phenomenology of extreme fluctuations of the \emph{instantaneous} drag and pointed out that the time scale associated with the development of such event is of the order of the sweeping time of the flow past the obstacle $\tau_c$.
We discussed previously the phenomenology of extreme fluctuations of the \emph{instantaneous} drag, and identified the sweeping time of the flow past the obstacle as the characteristic lifetime of these events.
%
In applications, this duration may be much smaller than the response time of the material structure subject to these fluctuations, justifying a practical interest in the averaged (in time) drag force.
% Consider for instance the interaction of a deformable structure with a turbulent flow: the typical response time may be much larger than the lifetime of drag fluctuations.
%
% definition
%
Therefore, a relevant observable is the \textit{time-averaged} drag defined as
\begin{equation}
\label{eq:def_time_averaged_drag}
F_T(t) = \frac{1}{T}\int_t^{t+T} f_d(t) \mathrm{d}t,
\end{equation}
where $f_d(t)$ denotes the instantaneous drag and $T$ is the investigated timescale (response time).
Relevant values of $T$ are application dependant.
We shall consider $T=10\tau_c$ in the following.
\EL{In that case, the PDF of $F_T$ is found nearly Gaussian as a consequence of the Central Limit theorem (see Fig.~\ref{fig}).}

%
During a time interval $[t;t+T]$, a fluctuation of $F_T(t)$ may be roughly viewed as the overall contribution of  $T / \tau_c$ independent fluctuations of the {instantaneous} drag $f_d$.
% phenomenology
%
% What is the phenomenology leading to extreme values of $F_T(t)$ ?
It is thus legitimate to ask whether a large value of the averaged drag results from a single outstanding fluctuation of the instantaneous drag (case (1)),  or from  an unusual succession of moderate positive fluctuations (case (2)).
%
%
%The averaged drag signal $\{F_T(t)\}_{0 \leq t \leq T_{tot}-T}$ is obtained by filtering in time the instantaneous drag signal $f_d(t)$.
In the same way as in section \ref{sec:extreme_extraction}, one can identify extreme fluctuations of $F_T$ exceeding some fixed threshold $a$, and sample a set of extreme events.
By taking $a=5.2\sigma_T$ with $\sigma_T$ being the standard deviation of $F_T$, $84$ independent extreme events have been selected. \EL{Again, this choice results from a compromise between the need to consider large deviations from the mean value and the requirement to sample a sufficient number of events for meaningful statistics. As a rule of thumb, the threshold has been set so that around one hundred events are selected.}

% neither case (1) nor case (2)
%
Fig.~\ref{fig:extreme_avg} displays the time-series $\{f_d(t)\}_{t^{\star} \leq t \leq t^{\star}+T}$ for several extreme fluctuations of $F_T$ occurring at $t=t^\star$.
\EL{Qualitatively, it is found that extreme fluctuations of the time-averaged drag can neither be reduced to case (1) nor case (2).
Indeed, both cases are %equally
featured in Fig.~\ref{fig:extreme_avg};
very large value of the averaged-drag appear to result from either a very large fluctuation, or a significant succession of moderate (positive) fluctuations of the instantaneous drag.}

\bigskip

\ZZ{The paragraph to justify equal probability of case (1) and case (2) is removed because it is too speculative (referee's remark)}

\bigskip

% theoretical explanation is removed because it is speculative

%This observation can be related to the exponential shape of the tail of the \ac{pdf} describing extreme positive drag fluctuations.
%Indeed, let $X$ be a random variable with a \ac{pdf} $\mathbb{P}(X)$ and a standard deviation $\sigma_X$.
%Considering an extreme positive value of $S_N=\sum_{n=1}^{N}X_n$, the probability $p_1$ (resp. $p_2$) of case (1) (resp. case (2)) writes
%\begin{equation}
%\label{eq:indep}
%p_{1}  \left(\sum_{n=1}^{N} X_n = S_N\right) \propto \mathbb{P} \left({S_N}/{N} \right)^{N} \quad \text{and} \quad p_{2}  \left(\sum_{n=1}^{N} X_n= S_N\right) \propto \mathbb{P}(S_N)
%\end{equation}
%If $\mathbb{P}$ has an exponential positive tail, \textit{i.e.} $\mathbb{P}(X=x) \underset{x \gg \sigma_X}{\propto} e^{-\lambda x}$, both cases (1) and (2) have equivalent probabilities provided that the average $a=S_N/N$ is very large:
%\begin{equation}
%\frac{p_{2}}{p_{1}} \underset{a\to \infty}{\to} \left(e^{\lambda a }\right)^{N} e^{-\lambda a N} = 1
%\label{eq:ratio_exp}
%\end{equation}
%The observation is therefore well supported theoretically.
%See~\cite{lestang2019JSTAT} for further details.

\section{Rare-events algorithms}
\label{sec:rare_events_algorithms}

According to Eq.~\eqref{eq:return_time}, the return time of fluctuations $f_d \geq a$ scales like
$r(a)\propto e^{l a}$, meaning that the computational cost required to sample events of amplitude $f_d \geq a$ through brute-force sampling diverges exponentially.
The direct sampling approach used in the previous sections to sample extremes of the drag is therefore not scalable and it would be difficult, if even feasible, to study rarer events or more complex flows.

In this section, we investigate the use of rare-events algorithms in order to sample extreme events for a
computational cost much lower than their return time.
Should this approach be successful, it would pave the way to the simulation of extreme events for flows of
greater complexity than the two-dimensional flow discussed in the previous sections.

We focus on two algorithms, representatives of two popular approaches in rare event simulation.
The first one is the \acl{ams} algorithm \citep{cerou_adaptive_2007} and builds on previous ideas about splitting approaches \citep{KahnHarris1951,glasserman_multilevel_1999}.
In recent years, it has allowed for the computation of rare events in problems involving a large number of degrees of freedom, such as molecular dynamics simulations \citep{aristoff_adaptive_2015,teo_adaptive_2016} or stochastic partial differential equations for the computation of rare trajectories in the Allen-Cahn equations \citep{rolland_computing_2016}. More recently, \EL{it} has been applied to rare events in stochastic models of wall-turbulence \citep{rolland_extremely_2018} and atmospheric dynamics \citep{bouchet2019rare}.
The second algorithm we consider in this section is the cloning algorithm \citep{giardina_direct_2006}, also known as population dynamics or interacting particles system algorithm.
This approach is very similar to the Diffusion Monte Carlo method used in computational chemistry, and has been introduced and refined in many variants.
Among them the cloning algorithm introduced by Giardina and Peliti \citep{giardina_direct_2006} was successfully applied to chaotic dynamical systems \citep{giardina_simulating_2011,Laffargue_2013}.
More recently it allowed the numerical simulation of extreme heat waves in a simplified modelling of the atmosphere \citep{ragone_computation_2018}.
This use represents a significant leap in the applicability of rare-event sampling algorithms to complex dynamical systems.
Along the same line, \EL{rare-event sampling algorithms are here applied outside of traditional applications by considering fluid-structure interaction in a turbulent flow}.

The \ac{ams} is based on the simulation of an ensemble of $N$ independent trajectories $\{\mathbf{x}_n(t)\}_{0\leq t \leq T_a}$ with $n=1 \cdots N$.
Each $\{\mathbf{x}_n(t)\}_{0\leq t \leq T_a}$ refers to a trajectory of duration $T_a$ in the phase space of the
system.
Following the initial simulation of this ensemble, trajectories are iteratively resampled in order to maximise the maximum value of a score function $\xi(\mathbf{x})$ along the trajectories.
The difficulty in applying the \ac{ams} lies in the choice of the score function, which can drastically
influence the effectiveness of the resampling in leading to extreme events \citep{rolland_statistical_2015}.
Mathematical analysis of the \ac{ams} provide proof of existence of an optimal score function, kown as the
committor function. However, with the exception of rare analytical cases, this function is unknown.
See appendix~\ref{app:AMS_on_OU} for a description of the \ac{ams} and its basic mathematical properties.

Similarly to the \ac{ams}, the cloning algorithm simulates an ensemble of trajectories of the system.
However, resampling steps are performed along the simulation of the ensemble.
Trajectories are first evolved independently from $t = 0$ to $t=\tau < T_a$ and then duplicated or pruned
according to the average of an observable $O(\mathbf{x})$ over the interval $[0;\tau]$.
This procedure is then repeated $n-1$ times over the intervals $[\tau, 2\tau], [2\tau, 3\tau]... [(n-1)\tau, n\tau = T_a]$.
Ultimately, the sampled trajectories are distributed according to a probability distribution that is tilted towards large values of the average observable:
\begin{align}
\mathbb{P}_{k}\left(\left\{ \mathbf{X}(t)\right\} _{0\leq t\leq T_{a}}=\left\{ \mathbf{x}(t)\right\} _{0\leq t\leq T_{a}}\right) &\underset{N\rightarrow\infty}{\sim} \frac{e^{k\int_{0}^{T_{a}}O(\mathbf{x}(t))dt}}{Z(k,T_a)}\mathbb{\mathbb{P}}_{0}\left(\left\{ \mathbf{X}(t)\right\} _{0\leq t\leq T_{a}}=\left\{ \mathbf{x}(t)\right\} _{0\leq t\leq T_{a}}\right),
\label{eq:Biased_Path_Approximation_main}
\end{align}
where
$\mathbb{P}_{0}\left(\left\{ \mathbf{X}(t)\right\} _{0\leq t\leq T_{a}} = \left\{ \mathbf{x}(t)\right\} _{0\leq t\leq T_{a}}\right)$ refers formally to the probability of observing the trajectory
$\left\{ \mathbf{x}(t)\right\} _{0\leq t\leq T_{a}}$ according to the unbiased statistics of the model.
See appendix~\ref{app:gktl_description} for a more detailed description of the cloning algorithm.

Both algorithms are similar in the way that they evolve an ensemble of trajectories, introducing selection
rules to bias the sampling towards extreme events.
In both cases, the output of the algorithm is an ensemble of $N$ trajectories that over-samples the extreme
events of interest.
Note that these trajectories are not statistically independent, however rigorous mathematical results link the statistics sampled by the algorithms to the unbiased statistics of the system, \textit{i.e.} the statistics sampled by a direct sampling approach \citep{DelMoral2013, brehier:hal-01142704}.

>>>>

\include{section_description_case_study}

\include{section_extremes_direct_sampling}

\include{rare_events_algorithms}
